---
title: "Overall analysis"
author: "Miao Zhang"
date: "2/6/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
load("all_environment.RData")
```

# 1. Overall kinematic analysis 
## 1.1. Gesture Duration
```{r gdur, fig.align="center"}
all_gdur_fig
```

Fig 1. displays the distribution of gesture duration data across 2 speakers in different conditions. I found a main effect of Information Structure on *gesture duration*: F(3, 286) = 8.60 (p < .001). *Across-accentuation*, longer movements were confirmed when comparing the maximally diverging focus structures, i.e., *background* and *contrastive focus*. The duration of opening gesture across the three different articulators increased on average 9.40ms. *Within-accentuation*, opening gestures were averagely 15.7ms longer in *contrastive focus* condition than *background focus* condition.

The main factor *Place of Articulation* also has an effect on *gesture duration* (F(2, 286) = 204.50 (p < .001), post-hoc: /k/ > /t/ > /p/).

The interaction between *place of articulation* and *information_structure* was confirmed significant (F(6, 286) = 2.56, p <. 05). While the greatest increase for *lower lip* is between *contrastive focus* and *background* (6.4ms longer), the largest differences for *tongue tip* and *tongue body* come from comparing *contrastive focus* and *broad focus* (26.2ms and 18ms longer, respectively). 

```{r gdur place:info_str}
gdur_info_place
```

Effects of *Footedness* on *gesture duration* failed to reach the level of significance. 

## 1.2. Maximum Displacement
```{r mdisp, fig.align="center"}
all_mdisp_fig
```

Fig 2. shows the distribution of maximum displacement of the opening gesture in different conditions. I observed larger movements across different *Information structures*: F(3, 286) = 6.82 (p < .001). *Across-accentuation*, the opening gestures are averagely 1.14mm larger in *contrastive focus* than in *background*. *Within-accentuation*, the displacement of gestures increased from *broad focus* to *contrastive focus* for an average of 1.37mm, from *narrow focus* to *contrastive focus* for an average of 0.85mm.

*Place of articulation* again has an effect on displacement (F(2, 286) = 208.58 (p < .001), post-hoc: /k/ > /t/ > /p/).

Comparing the two *footedness* condition, the movement under footed conditions overall is 0.48mm larger than unfooted conditions (F(1, 286) = 4.36 (p < .05)).

The interaction of *place* and *footedness* reached the level of significance (F(2, 286) = 5.075 (p < .01)).

```{r mdisp place:footedness}
mdisp_place_foot
```

As can be seen from the table, whereas for *lower lip* and *tongue body*, the articulators moved larger from unfooted condition to footed condition, *tongue tip* moved smaller.

## 1.3. Peak Velocity
```{r pvel, fig.align="center"}
all_pvel_fig
```

No significant differeces due to *information structure* was confirmed but it was statistically close (F(3, 286) = 2.26, p < .1). On the other hand, *footedness* (F(1, 286) = 4.09 (p < .05); post-hoc: 8.23mm/s faster when footed than unfooted) and *place of articulation* (F(2, 286) = 25.1498 (p < .05); post-hoc: /k/ > /t/ > /p/) both have a main effect on how fast the articulators move. Interactions between the three variables were not significant for peak velocity. 

## 1.4. Time-to-Peak Velocity
```{r t2pvel, fig.align="center"}
all_t2pvel_fig
```

An effect of *Information structureand* on *time-to-peak velocity* was observed (F(3, 286) = 6.12 (p < .05)). *Across-accentuatoin*, there is a less stiff movement comparing *contrastive focus* to *background* (6.34ms later). *Within-accentuation* time-to-peak velocity is 9.83ms later in *contrastive focus* than in *broad focus*. Though not significant, the difference between contrastive focus and narrow focus of 5.78ms is just above significant level (p-adjusted = 0.06).

The *place of articulation* again is significant (F(2, 286) = 59.41 ( p < .05); post-hoc: /k/ > /t/ > /p/).

There was also an interaction between *place of articulation* and *information structure* (F(6, 286) = 3.34, p < .01). The greatest modification in articulation for lower lip comes from comparing narrow and background context (2.3ms later in narrow focus), while for tongue tip and tongue body, the greatest increase was found in comparing contrastive and broad focus (23.9ms and 12.9ms respectively).

```{r t2pvel place:footedness}
t2pvel_place_info
```

# 2. By-speaker analysis
There is a huge inter-speaker variation observed from the data. A series of two-way ANOVAs taking *information structure* and *footedness* as the independent variables was run by each speaker for each kinematic measurement. The table below summarises the result of the analysis by speakers. 

to be continued...